### 2020-06-30 chapters/background.tex
Images for mechant1 features

Batch processing is the processing of data grouped by batches. A batch is a collection of data points that have been grouped by certain criteria. Examples of such criteria could be \textit{"all the transactions from the past two months"}, \textit{"the last million transactions"} or even \textit{"all the transactions made by an user until now"}. Batch data sets are known as static data \cite{Martin-Batch-Defin}. An example application of such a technique would be processing all the keywords searched for in the google search engine, and then grouping the results by keyword to count how many times a keyword was used in searches (for trend analysis purposes). Batch processing is the oldest paradigm between batch and stream processing. Historically speaking, Apache Hadoop \cite{borthakur2007hadoop} \cite{Hadoop} \cite{ApacheHadoop} was one of the first widely known frameworks for large-scale batch processing, created in 2005. Hadoop uses a distributed computation framework or programming model called MapReduce \cite{MapReduce}. This paradigm applies a map function that processes key/value pairs and generates sets of intermediate ones, ultimately merging all intermediate values
using a reducer function. To do this they read and write from disk on intermediate steps, which makes it inefficient for applications that reuse working sets of data across multiple parallel operations (\textit{e.g.} iterative Machine Learning algorithms). Nearly a decade later, Apache Spark \cite{ApacheSpark} \cite{Spark} was released.
Spark retains the scalability and fault tolerance of MapReduce but introduces Resilient Distributed Datasets (RDDs) \cite{SparkRDDs}. An RDD is a read-only collection of objects partitioned across a set of machines that can be rebuilt if a partition is lost.
Additionally, Spark processes data in random access memory (RAM), while Hadoop MapReduce persists data to the disk after a map or reduce action. Thus, Spark achieves lower processing times.

On the other hand, stream processing refers to the processing of data streams. A data stream is essentially a continuous or unbounded data set. This data needs to be processed sequentially and incrementally on a event-by-event basis or over sliding windows. Streaming data includes a wide variety of data such as logs generated by customers using mobile or web applications, sensor data from Internet-of-Things (IoT) devices or credit card transactions. This is in contrast to batch processing which works on a finite data set. This distinction is important since some aggregations are very hard and even impossible to compute with 100\% accuracy for several use cases in an efficient manner on unbounded data sets. An example of such is counting distinct elements that flow through the data stream. An immediate solution would be to create a regular set of items or strings and have its length be the number of distinct elements processed. However, this approach is not very efficient and has a linear memory usage growth regarding the number of distinct elements. Therefore, as the number of distinct elements approaches large orders of magnitude we run out of memory.

@@

Aggregations are present in most programming languages and frameworks. Quoting from the Java programming language documentation on the \textit{Stream.reduce()} method \cite{Java13StreamReduce}: 
\textit{"A reduction operation (also called a fold) takes a sequence of input elements and combines them into a single summary result by repeated application of a combining operation"}. Similarly, the MapReduce \cite{MapReduce} paradigm consists of first applying \textit{map} steps to the collection of data and then a \textit{reduce} one. \textit{Map} tasks will transform each item as desired and the \textit{reduce} task will merge the collection into a single value. Reduce methods make use of a \textit{combine} function and recursively apply it. The \textit{combine} function takes the current aggregation value and combines it with the next item in the set, returning the new aggregated state. Doing this recursively for each item in the collection results in a single value: the aggregate value. The combine function is here on forward represented by $\oplus$ and the identity or neutral element of the aggregation by $\overline{\theta}$.

>>>>

Batch processing is the processing of data grouped by batches. A batch is a collection of data points that have been grouped by certain criteria. Examples of such criteria could be \textit{"all the transactions from the past two months"}, \textit{"the last million transactions"} or even \textit{"all the transactions made by an user until now"}. Batch data sets are known as static data \cite{Martin-Batch-Defin}. An example application of such a technique would be processing all the keywords searched for in the google search engine, and then grouping the results by keyword to count how many times a keyword was used in searches (for trend analysis purposes). Batch processing is the oldest paradigm between batch and stream processing. Historically speaking, Apache Hadoop \cite{borthakur2007hadoop} \cite{Hadoop} \cite{ApacheHadoop} was one of the first widely known frameworks for large-scale batch processing, created in 2005. Hadoop uses a programming model called MapReduce \cite{MapReduce}. This paradigm applies a map function that processes key/value pairs and generates sets of intermediate ones, ultimately merging all intermediate values
using a reducer function. To do this, the distributed computation framework reads and writes from disk on intermediate steps, which makes it inefficient for applications that reuse working sets of data across multiple parallel operations (\textit{e.g.} iterative Machine Learning algorithms). Nearly a decade later, Apache Spark \cite{ApacheSpark} \cite{Spark} was released as the industry accepted successor.
Spark retains the scalability and fault tolerance of MapReduce but introduces Resilient Distributed Datasets (RDDs) \cite{SparkRDDs}. An RDD is a read-only collection of objects partitioned across a set of machines that can be rebuilt if a partition is lost.
Additionally, Spark processes data in random access memory (RAM), while Hadoop MapReduce persists data to the disk after a map or reduce action. Thus, Spark achieves lower processing times.

Stream procesisng, on the other hand, refers to the processing of data streams. A data stream is essentially a continuous or unbounded data set. This data needs to be processed sequentially and incrementally on a event-by-event basis or over sliding windows. Streaming data varies widely, from logs generated by customers using mobile or web applications, sensor data from Internet-of-Things (IoT) devices to credit card transactions. This is in contrast to batch processing which works on a finite data set. This distinction is important since some aggregations are very hard and even impossible to compute with 100\% accuracy for several use cases in an efficient manner on unbounded data sets. An example of such is counting distinct elements that flow through the data stream. An immediate solution would be to create a regular set of items or strings and have its length be the number of distinct elements processed. However, this approach is not very efficient and has a linear memory usage growth regarding the number of distinct elements. Therefore, as the number of distinct elements approaches large orders of magnitude we run out of memory.

@@

Aggregations are present in most programming languages and frameworks. Quoting from the Java programming language documentation on the \textit{Stream.reduce()} method \cite{Java13StreamReduce}: 
\textit{"A reduction operation (also called a fold) takes a sequence of input elements and combines them into a single summary result by repeated application of a combining operation"}. Similarly, the MapReduce \cite{MapReduce} paradigm consists of first applying \textit{map} steps to the collection of data and then a \textit{reduce} one. \textit{Map} tasks will transform each item as desired and the \textit{reduce} task will merge the collection into a single value. Reduce methods make use of a \textit{combine} function and recursively apply it. The \textit{combine} function takes the current aggregation value and combines it with the next item in the set, returning the new aggregated state. Doing this recursively for each item in the collection results in a single value: the aggregate value. The combine function is here on forward represented as $\oplus$ and the identity or neutral element of the aggregation as $\overline{\theta}$.

### 2020-06-30 chapters/introduction.tex
problem statement
Images for mechant1 features

\section{Context} \label{sec:context}
In the past decade, applications have become increasingly data-driven, placing data at the center of application design. For different use cases, the data is processed in different ways for different purposes. For instance, e-commerce platforms such as Alibaba\footnote{https://alibaba.com} need to process a large number of daily transactions while ensuring that sales run smoothly and that products are delivered to customers' homes. Streaming services, like YouTube\footnote{https://youtube.com} and Netflix\footnote{https://netflix.com}, try to guarantee that media content reaches up to millions of users simultaneously. Social networks are also responsible for generating large volumes of data,Twitter\footnote{https://twitter.com} for example, generates more than 500 million \textit{tweets} (posts in Twitter) per day\footnote{https://blog.twitter.com/engineering/en\_us/a/2013/new-tweets-per-second-record-and-how.html}. \textit{Tweets} may contain text, media content or both. Cybersecurity applications are yet another example of time-sensitive, data-driven applications. These need to monitor user accesses and their actions in the network, where timely detection of intruders is critical to prevent them from tampering with the protected system.

@@

As mentioned, the processing of data can be divided into two main categories: batch and stream processing. However, due to the increasing volume of data and the need for timely processing to allow companies to react to changing conditions in real-time, the demand for stream processing systems is increasing. Unlike batch processing of static data sets, where assumptions on the underlying data distributions can be made, streaming data is known for being non-stationary \cite{Gama-Knowledge-Discovery}. Furthermore, in the streaming paradigm, the value of the produced information lies in its recency \cite{Kolajo-Big-data-stream-SLR}. \textit{Recency} is measured under different scales for different use cases. For instance, consider the monitoring of geological data to forecast possible natural disasters such as earthquakes versus the monitoring of a computer network for intruder detection. In the former case, after the forecast of a future earthquake, governmental authorities need a couple of days to launch an evacuation plan and keep the population safe. However, in the intruder detection scenario, the decision of removing access to a user must be done as soon as possible, preferably in a few seconds. Hence, while in the first scenario information retrieved within a day would still be recent, in the second scenario, information is considered recent and relevant if delivered within seconds (or even milliseconds).

In either processing scenario, we usually want to apply operations on data, for example, to compute the maximum value, an average or a count of distinct elements. Operations that produce a single result when applied to data sets are commonly known as aggregation operations. In stream processing, we apply these aggregations to data streams to obtain valuable information. However, since a data stream is, effectively speaking, an unbounded data set, if our aggregation requires a finite domain, there is the need to apply windowing techniques. A thorough analysis of stream windowing techniques is done in the Background Section of this document but for now it suffices to say that windowing transforms an unbounded data stream into a finite data set by defining a time or tuple-based window size --- \textit{i.e.} events from the last three days or the last million events, respectively.

@@

Many market solutions monitor streams of credit card transactions with the intent of fraud detection. In the context of this Thesis, we work side by side with Feedzai\footnote{https://feedzai.com}. Feedzai offers fraud prevention solutions for the previously presented use case of monitoring a continuous stream of credit card transactions and detecting fraud. We exploit multiple real-life data sets that are provided from the financial fraud space.

Feedzai's solution for fraud detection consists of a workflow of different components. Incoming events are fed into this workflow and processed accordingly. Workflow components can be sets of user-defined rules and/or Machine Learning (ML) models that output scores reflecting their belief on whether a transaction is fraudulent or not. ML models are configured and trained,  before deployment, often with assumptions on the statistical distributions of future incoming data. Thus their model performance depends on how well those assumptions hold for future data. However, fraud patterns are not static over time, but seasonal --- \textit{e.g.} Black Fridays, holiday shopping and product launches --- and evolutionary --- \textit{e.g.} new fraud attack methods by fraudsters. Due to these shifts in data patterns and distributions, the ML models' performance can deteriorate over periods of time. In the Machine Learning field, this performance decay caused by data pattern shifts is also known as concept drift.

@@

The objective of this Thesis is to build a lightweight real-time system that monitors a stream of high volumes of high velocity, highly skewed and seasonal data and detects pattern shifts relative to a reference period.

>>>>

\section{Context} \label{sec:context}
In the past decade, applications have become increasingly data-driven, placing data at the center of application design. For different use cases, the data is processed in different ways for different purposes. For instance, e-commerce platforms such as Alibaba\footnote{https://alibaba.com} need to process a large number of daily transactions while ensuring that sales run smoothly and that products are delivered to customers' homes. Streaming services, like YouTube\footnote{https://youtube.com} and Netflix\footnote{https://netflix.com}, try to guarantee that media content reaches up to millions of users simultaneously. Social networks are responsible for generating large volumes of data of which,  Twitter\footnote{https://twitter.com} for example, generates more than 500 million \textit{tweets} (posts in Twitter) per day\footnote{https://blog.twitter.com/engineering/en\_us/a/2013/new-tweets-per-second-record-and-how.html}. \textit{Tweets} may contain text, media content or both. Cybersecurity applications are yet another example of time-sensitive, data-driven applications. The entities that work in these use-cases need to monitor user accesses and their actions in the network, where timely detection of intruders is critical to prevent them from tampering with the underlying system.

@@

As mentioned, the processing of data can be divided into two main categories: batch and stream processing. However, due to the increasing volume of data and the need for timely processing to allow companies to react to changing conditions in real-time, the demand for stream processing systems is increasing. Unlike batch processing of static data sets, where assumptions on the underlying data distributions can be made, streaming data is known for being non-stationary \cite{Gama-Knowledge-Discovery} where the value of the produced information lies in its recency \cite{Kolajo-Big-data-stream-SLR}. \textit{Recency} is measured under different scales for different use cases. For instance, consider the monitoring of geological data to forecast possible natural disasters such as earthquakes versus the monitoring of a computer network for intruder detection. In the former case, after the forecast of a future earthquake, governmental authorities need a couple of days to launch an evacuation plan and keep the population safe. However, in the intruder detection scenario, the decision of removing access to a user must be done as soon as possible, preferably in a few seconds. Hence, while in the first scenario information retrieved within a day would still be recent, in the second scenario, information is considered recent and relevant if delivered within seconds (or even milliseconds).

In either processing scenario, we usually want to apply operations on data, for example, to compute the maximum value, an average or a count of distinct elements. Operations that produce a single result when applied to data sets are commonly known as aggregation operations. In stream processing, we apply these aggregations to data streams to obtain valuable information. However, since a data stream is, effectively speaking, an unbounded data set, if our aggregation requires a finite domain, there is the need to apply windowing techniques. A thorough analysis of stream windowing techniques is done in the Background Section of this document but for now it suffices to say that windowing transforms an unbounded data stream into a finite data set typically by defining a time or tuple-based window size --- \textit{i.e.} events from the last three days or the last million events, respectively.

@@

Many market solutions monitor streams of credit card transactions with the intent of fraud detection. In the context of this Thesis, we work side by side with Feedzai\footnote{https://feedzai.com}. Feedzai offers fraud prevention solutions for the previously presented use case of monitoring a continuous stream of credit card transactions and detecting fraud. We exploit real-life data sets that are provided from the financial fraud space.

Feedzai's solution for fraud detection consists of a workflow of different components. Incoming events are fed into this workflow and processed accordingly. Workflow components can be sets of user-defined rules and/or Machine Learning (ML) models that output scores reflecting their belief on whether a transaction is fraudulent or not. ML models are configured and trained,  before deployment, often with assumptions on the statistical distributions of future incoming data. Thus their model performance depends on how well those assumptions hold for future data. However, fraud patterns are not static over time, but seasonal --- \textit{e.g.} Black Fridays, holiday shopping and product launches --- and evolutionary --- \textit{e.g.} new fraud attack methods by fraudsters. Due to these shifts in data patterns and distributions, the ML models' performance can deteriorate over periods of time. In the Machine Learning field, this performance decay caused by data pattern shifts is also known as concept drift \cite{drift} \cite{gama-drift}.

@@

Thus the objective of this Thesis is to build a lightweight real-time system that monitors a stream of high volumes of high velocity, highly skewed and seasonal data and detects pattern shifts relative to a reference period.

### 2020-06-30 chapters/my-work.tex
Nearly done with experiments
Images for mechant1 features

One aggregation that encodes distributions properly are histograms. Histograms are a representation of the distribution of numerical data thus being ideal to encode the distribution of feature values, satisfying \textit{(c)}. Histograms require storing a fixed-size list of bins and corresponding counts per bin regardless of the volume of data, satisfying property \textit{(a)}. Adding or removing an element to the histogram aggregation is tantamount to determining the correct bin and incrementing or decrementing the associated count, respectively, thus satisfying property \textit{(b)}. Recall Section \ref{sec:back-swag-algs} where we discussed generic sliding window aggregation algorithms and remember the Subtract-On-Evict (SOE) algorithm (Algorithm \ref{pseudo:soe}). Computing a histogram aggregation over a sliding window of size $W$ using SOE implies storing all $W$ elements in memory to know the element to remove after a new one is inserted. This violates our constant in-memory property \textit{(a)} which is why we chose to use the Exponential Moving Averages (EMAs) analyzed in Section \ref{sec:emas}. EMA-like histograms computed over sliding windows do not require the storage of window contents. An EMA-like histogram is \textit{(a)} constant in memory, \textit{(b)} maintained in real-time and \textit{(c)} encodes the distribution of each feature values. EMA-like histograms are further explained in Section \ref{sec:ema-hist}.

@@

\subsubsection*{Batch phase artifacts}
By the end of the batch phase, we should have, for each feature: 
\begin{itemize}
    \item a reference histogram
    \item a list of distance values that represent its distribution
    \item the last sample's histogram, to be used as burn-in period for the target sliding window histogram
\end{itemize}

>>>>

One aggregation that encodes distributions properly are histograms. Histograms are a representation of the distribution of numerical data thus being ideal to encode the distribution of feature values, satisfying \textit{(c)}. Histograms require storing a fixed-size list of bins and corresponding counts per bin regardless of the volume of data, satisfying property \textit{(a)}. Adding or removing an element to the histogram aggregation is tantamount to determining the correct bin and incrementing or decrementing the associated count, respectively, thus satisfying property \textit{(b)}. Recall Section \ref{sec:back-swag-algs} where we discussed generic sliding window aggregation algorithms and remember the Subtract-On-Evict (SOE) algorithm (Algorithm \ref{pseudo:soe}). Computing a histogram aggregation over a sliding window of size $W$ using SOE implies storing all $W$ elements in memory to know the element to remove after a new one is inserted. This violates our constant in-memory property \textit{(a)} which is why we chose to use the Exponential Moving Averages (EMAs) analyzed in Section \ref{sec:emas}. EMA-like histograms computed over sliding windows do not require the storage of window contents. An EMA-like histogram is \textit{(a)} constant in memory, \textit{(b)} maintained in real-time and \textit{(c)} encodes the distribution of each feature values. These EMA-like histograms are further explained in Section \ref{sec:ema-hist}.

@@

\subsubsection*{Batch phase artifacts} \label{sec:batch-artifacts-summary}
By the end of the batch phase, we should have, for each feature: 
\begin{itemize}
    \item a reference histogram
    \item a list of distance values that represent its distribution
    \item the last sample's histogram, to be used as burn-in period for the target sliding window histogram
\end{itemize}

### 2020-06-30 chapters/problem-statement.tex
Images for mechant1 features

\section{The problem}
Many real-time stream monitoring systems are static once deployed in a production environment. Engineers of said systems configure them under the assumption that future data flowing through the system follows roughly the same distribution as previously seen data. They do so consciously, to the best of their knowledge and available tools, keeping in mind that in the future they may need to reconfigure the system. Thus, even though the initial configuration of the system may be one of the best fits, over time, due to data pattern shifts, the initially deployed static system's performance gradually deteriorates. The problem at hand is determining when to reconfigure the system, whatever it may be (\textit{e.g.} a Machine Learning model), based on streaming analysis.

\section{State of the art issues}
Aggregations, specifically sliding window ones, were introduced in Chapter \ref{chap:background} along with several windowing techniques required to monitor streaming data patterns in real-time. In Chapter \ref{chap:sota} we presented state of the art sliding window aggregation algorithms as well as outlier detection methods. However, the reviewed methods were deemed unfit for our use-case because none of them simultaneously showed all of the below characteristics:

@@

    \item \textbf{Fixed reference period:} the reference window must be kept the same throughout run-time, the opposite of an online learning algorithm

@@

    \item \textbf{Explainable alerts:} have some measure of divergence to understand why an alert was raised
\end{itemize}

\section{Proposal}
We devise a two-phased (batch and streaming phases) and two-windowed (reference and target sliding windows) method. We devise a sliding window approximate histogram aggregation that encodes data distributions and provide a method for incrementally maintaining it with constant time and space complexity for every growing volumes of data, ideal for the streaming scenario. Given a user-defined reference period and a stream of data to process, our method guarantees to monitor each event's fields and alert if they are considered divergent from their reference distribution. We provide a measure of how divergent each field (or feature) is. Based on the number of divergent features and the corresponding divergence measure, a system administrator has an easier time determining when to reconfigure the system.

@@

\begin{enumerate}[label=\textbf{(RQ\arabic*)}]
    \item What aggregation can we use to properly encode data distributions?
    \item How can we maintain said aggregation in a sliding window fashion in real-time with constant memory usage?
    \item How can we measure divergence between the reference and sliding window aggregations?
    \item Based on a divergence measure, when should we raise an alert?
\end{enumerate}

>>>>

\section{The problem}
Many real-time stream monitoring systems are static once deployed in a production environment. Engineers of said systems configure them under the assumption that future data flowing through the system follows roughly the same distribution as previously seen data. They do so consciously, to the best of their knowledge and available tools, keeping in mind that in the future they may need to reconfigure the system. Thus, even though the initial configuration of the system may be one of the best fits, over time, due to data pattern shifts, the initially deployed static system's performance gradually deteriorates. The problem at hand is to determine when to reconfigure the system, \textit{e.g.} a Machine Learning model, based on streaming analysis.

\section{State of the art issues}
Aggregations, specifically sliding window ones, were introduced in Chapter \ref{chap:background} along with several windowing techniques required to monitor streaming data patterns in real-time. In Chapter \ref{chap:sota} we presented state of the art sliding window aggregation algorithms as well as outlier detection methods. However, the reviewed methods were deemed unfit for our use-case because none of them simultaneously showed all of the desired characteristics, namely:

@@

    \item \textbf{Fixed reference period:} the reference window must be kept the same throughout run-time

@@

    \item \textbf{Explainable alerts:} have an interpretable measure of divergence to understand why an alert was raised
\end{itemize}

\section{Proposal}
We devise a two-phased (batch and streaming phases) and two-windowed (reference and target sliding windows) method. We devise a sliding window approximate histogram aggregation that encodes data distributions and provide a method for incrementally maintaining it with constant time and space complexity for every growing volumes of data, ideal for the streaming scenario. Given a user-defined reference period and a stream of data to process, our method guarantees to monitor each event's fields and alert if they are considered divergent relative to their reference distribution. We provide a measure of how divergent each field (or feature) is. Based on the number of divergent features and the corresponding divergence measure, a system administrator has an easier time determining when to reconfigure the system.

@@

\begin{enumerate}[leftmargin=1.5cm, label=\textbf{(RQ\arabic*)}]
    \item What aggregation can we use to properly encode data distributions?
    \item How can we maintain such an aggregation in a sliding window fashion in real-time with constant memory usage?
    \item How can we measure divergence between the reference and sliding window aggregations?
    \item Based on a divergence measure, when should we raise an alert?
\end{enumerate}

%Assumptions section
%
%
%
%
%

### 2020-06-30 chapters/validation.tex
Images for mechant1 features

In Section \ref{sec:ema-hist}, we mentioned that we choose to discard events whose contribution is less than 6\%. We set a half-life $n_{1/2}$ value and then multiply it by a constant factor of 4 to get the size of the samples to make (a larger sample would contain events whose contribution was equal or less than 6\%).

@@

Feedzai provides a fraud detection solution for this merchant and had stored two contiguous in time datasets of credit card transactions that combined covered a period of 1 year and 9 months. We used the older one as the reference dataset and the other one as the target one. Both datasets were already feature-engineered \cite{Domingos-ML-Feat-Eng} and had a total of 183 numerical plus categorical features. Table \ref{tbl:merchant1-datasets-summary} shows summary statistics for both datasets.
\begin{table}[!htb]
    \begin{center}
        \begin{tabular}{|c|c|c|c|}
        \hline
        \textbf{Dataset} & \textbf{From}              & \textbf{To}           & \multicolumn{1}{l|}{\textbf{Transactions}} \\ \hline
        Reference        & Tuesday, September 5, 2017 & Friday, June 29, 2018 & 1,604,509                                  \\ \hline
        Target           & Friday, June 29, 2018      & Sunday, June 30, 2019 & 4,032,505                                  \\ \hline
        \end{tabular}
        \caption{Summary statistics for reference and target merchant datasets}
        \label{tbl:merchant1-datasets-summary}
    \end{center}
\end{table}

In the batch phase, we used 100 Spark executors, each with 12G of memory. We determined that one week's worth of data for this merchant was on average equivalent to 37,803 events. We set the half-life to be $n_{1/2}=37,803$ and discard events whose contribution is 6\% or less, which is equivalent to using windows of data of four times the half-life (as detailed in Section \ref{sec:ema-hist}). Hence, our samples' windows had 151,212 events. Using Equation \ref{eq:optimal-n-samples}, we determined that for 183 features, a FWER of 1\% and $\gamma$ (as defined in Section \ref{sec:nsamples}) of 10\% we needed to make a minimum of 42,137 samples. We conclude the batch analysis in 5 hours and 34 minutes.

After the batch phase, we had the necessary inputs for the streaming phase (detailed in Section \ref{sec:batch-artifacts-summary}), namely, for each feature, the reference histograms, the list of observed divergence values and the last sample's histogram. We then used the target dataset, simulating a stream of ordered events. Our monitoring system processed event by event and reported divergent features and associated p-values. We performed our multiple hypotheses with the Holm-Bonferroni correction (Section \ref{sec:holmbonferroni}) for every 1000 events. It took us 382,363 milliseconds to process the 4,032,505 transactions, amounting to a throughput of 10,546 transactions per second (TPS).

Our system produced a large list of alerts. We had alerts from nearly the beginning of the target period until the end. To try and make sense of these alerts, we questioned Feedzai's data scientists that worked directly with these data sources. We found out that this merchant and in particular these datasets were known to suffer a lot from concept drift. Additionally, the data scientists performed some data preprocessing and cleaning, removing some of the missing values, but only on the reference dataset, which further augmented the differences between the target and reference datasets. These seemed like plausible reasons for the large number of alarms generated.

%x1 = completion_rate_per_card_72h
We set out to validate this hypothesis. Due to the large number of features (183) we only show the plots and visualizations drawn for one. Furthermore, we refer to features with anonymized identifiers, like our example feature, x1. Figure \ref{fig:merchant-x1-corrected-pvalues} shows the corrected p-values (Formula \ref{for:corrected-pvalue}) for feature x1 at for each event of the time-series and Figure \ref{fig:merchant-x1-corrected-pvalues-zoom} provides a closer look on the region where the p-values drop below the threshold (FWER = 1\%).
\begin{center}
\begin{minipage}{.5\textwidth}
  \centering
  \includegraphics[width=1\linewidth]{figures/merchant-x1-corrected-pvalues.png}
  \captionof{figure}{x1 corrected p-values and threshold}
  \label{fig:merchant-x1-corrected-pvalues}
\end{minipage}%
\begin{minipage}{.5\textwidth}
  \centering
  \includegraphics[width=1\linewidth]{figures/merchant-x1-corrected-pvalues-zoom.png}
  \captionof{figure}{x1 zoomed in corrected p-values and threshold}
  \label{fig:merchant-x1-corrected-pvalues-zoom}
\end{minipage}
\end{center}
We see that this feature drops below the alarm threshold and never rises above it, effectively being considered divergent for the rest of the target period. Furthermore, it reaches p-values very close to zero and eventually becomes zero, throughout the remaining target period. Is such divergence possible? To find out, let us plot the reference and target time-series for feature x1. Figure \ref{fig:merchant-x1-reference} shows the reference time-series plot and Figure \ref{fig:merchant-x1-target} shows the target time-series for feature x1.
\begin{center}
\begin{minipage}{.5\textwidth}
  \centering
  \includegraphics[width=1\linewidth]{figures/merchant-x1-reference.png}
  \captionof{figure}{x1 reference time-series}
  \label{fig:merchant-x1-reference}
\end{minipage}%
\begin{minipage}{.5\textwidth}
  \centering
  \includegraphics[width=1\linewidth]{figures/merchant-x1-target.png}
  \captionof{figure}{x1 target time-series}
  \label{fig:merchant-x1-target}
\end{minipage}
\end{center}
It became clear to us why this feature and many others were being alerted from the start of the dataset until the end. The domain of values is significantly different. The reference period was cleaned of missing values, usually represented by -1. Hence, our reference was built expecting values from 0 to 1 with a certain distribution and in the stream analysis, we found a large volume of missing values or -1. This explains the high number of alerts.

A follow-up question that is raised is if we should include some mechanism of dealing with missing values in our method. We discard this idea and defend that we should not: missing values are indeed something we want to detect. For instance, if a certain entity responsible for reporting one feature stops working, we want to know that. This scenario is very likely to happen in a distributed Internet-of-Things (IoT) system with hundreds of devices transmitting data.

From this experiment, we conclude that for high concept drift scenarios the system works properly, and alerts all divergent features.

\subsection{Similar reference and target periods}
In this next experiment, we set out to test our system in a low drift scenario,...

\textcolor{red}{SPLIT TARGET PERIOD IN TWO, USING SAME PERIOD GUARANTEES SIMILAR PERIODS? BUT MAYBE IT DOES NOT (LOOK AT x1 TARGET TIMESERIES, 1/3 has no MISSING VALUES AS WELL. STILL VERY DIFFERENT FROM 2/3)}.

>>>>

In Section \ref{sec:ema-hist}, we mentioned that we choose to discard events whose contribution is less than 6\%. We set a half-life $n_{1/2}$ value and then multiply it by a constant factor of 4 to get the size of the samples to make (a larger sample would contain events whose contribution was equal to or less than 6\%).

@@

Feedzai provides a fraud detection solution for this merchant and had stored two contiguous in time datasets of credit card transactions that combined covered a period of 1 year and 9 months. We used the older one as the reference dataset and the more recent one as the target one. Both datasets were already feature-engineered \cite{Domingos-ML-Feat-Eng} and had a total of 183 numerical plus categorical features. Table \ref{tbl:merchant1-datasets-summary} shows summary statistics for both datasets.
\begin{table}[!htb]
    \begin{center}
        \begin{tabular}{|c|c|c|c|}
        \hline
        \textbf{Dataset} & \textbf{From}              & \textbf{To}           & \multicolumn{1}{l|}{\textbf{Transactions}} \\ \hline
        Reference        & Tuesday, September 5, 2017 & Friday, June 29, 2018 & 1,604,509                                  \\ \hline
        Target           & Friday, June 29, 2018      & Sunday, June 30, 2019 & 4,032,505                                  \\ \hline
        \end{tabular}
        \caption{Summary statistics for reference and target merchant datasets}
        \label{tbl:merchant1-datasets-summary}
    \end{center}
\end{table}

In the batch phase, we used 100 Spark executors, each with 12G of memory. We determined that one week's worth of data for this merchant was on average equivalent to 37,803 events. We set the half-life to be $n_{1/2}=37,803$ and discard events whose contribution is 6\% or less, which is equivalent to using windows of data of four times the half-life (as detailed in Section \ref{sec:ema-hist}). Hence, our samples' windows had 151,212 events. Using Equation \ref{eq:optimal-n-samples}, we determined that for 183 features, a FWER of 1\% and $\gamma$ (as defined in Section \ref{sec:nsamples}) of 10\% we needed to make a minimum of 42,137 samples, concluding the batch analysis in 5 hours and 34 minutes.

After the batch phase, we had the necessary inputs for the streaming phase (detailed in Section \ref{sec:batch-artifacts-summary}), namely, for each feature, the reference histograms, the list of observed divergence values and the last sample's histogram. We then used the target dataset, simulating a stream of ordered events. Our monitoring system processed the dataset event by event in sequential, time-ordered fashion, emulating a stream computation and reported divergent features and associated p-values. We performed our multiple hypotheses with the Holm-Bonferroni correction (Section \ref{sec:holmbonferroni}) for every 1000 events. It took us 382,363 milliseconds (roughly 6 minutes) to process the 4,032,505 transactions, amounting to a throughput of 10,546 transactions per second (TPS).

\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.5]{figures/merchant-x1-reference.png}
      \caption{Merchant feature x1 reference time-series}
      \label{fig:merchant-x1-reference}
    \end{center}
\end{figure}

\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.5]{figures/merchant-x1-target.png}
      \caption{Merchant feature x1 target time-series}
      \label{fig:merchant-x1-target}
    \end{center}
\end{figure}

\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.5]{figures/merchant-x2-reference.png}
      \caption{Merchant feature x2 reference time-series}
      \label{fig:merchant-x2-reference}
    \end{center}
\end{figure}

\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.5]{figures/merchant-x2-target.png}
      \caption{Merchant feature x2 target time-series}
      \label{fig:merchant-x2-target}
    \end{center}
\end{figure}

\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.5]{figures/merchant-x3-reference.png}
      \caption{Merchant feature x3 reference time-series}
      \label{fig:merchant-x3-reference}
    \end{center}
\end{figure}

\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.5]{figures/merchant-x3-target.png}
      \caption{Merchant feature x3 target time-series}
      \label{fig:merchant-x3-target}
    \end{center}
\end{figure}

\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.5]{figures/merchant-x4-reference.png}
      \caption{Merchant feature x4 reference time-series}
      \label{fig:merchant-x4-reference}
    \end{center}
\end{figure}

\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.5]{figures/merchant-x4-target.png}
      \caption{Merchant feature x4 target time-series}
      \label{fig:merchant-x4-target}
    \end{center}
\end{figure}

\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.5]{figures/merchant-x5-reference.png}
      \caption{Merchant feature x5 reference time-series}
      \label{fig:merchant-x5-reference}
    \end{center}
\end{figure}

\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.5]{figures/merchant-x5-target.png}
      \caption{Merchant feature x5 target time-series}
      \label{fig:merchant-x5-target}
    \end{center}
\end{figure}

%This merchant and in particular these datasets were known to suffer a lot from concept drift. Our system produced a large list of alerts. We had alerts from nearly the beginning of the target period until the end.

%low drift:
    %x1 - max_amt_rate_per_location_72h
    %x2 - num_preauths_per_card_1w
    %x3 - user_avg_amount_completions_1h
    %x4 - completions_rate_per_location_zip3_72h

%high drift:
    %x5 - completion_rate_per_card_72h

%Due to the large number of features (183) we only show the plots and visualizations drawn for some. Furthermore, we refer to features with anonymized identifiers, like our example feature, x1. Figure \ref{fig:merchant-x1-corrected-pvalues} shows the corrected p-values (Formula \ref{for:corrected-pvalue}) for feature x1 at for each event of the time-series and Figure \ref{fig:merchant-x1-corrected-pvalues-zoom} provides a closer look on the region where the p-values drop below the threshold (FWER = 1\%).

%We see that this feature drops below the alarm threshold and never rises above it, effectively being considered divergent for the rest of the target period. Furthermore, it reaches p-values very close to zero and eventually becomes zero, throughout the remaining target period. Is such divergence possible? To find out, let us plot the reference and target time-series for feature x1. Figure \ref{fig:merchant-x1-reference} shows the reference time-series plot and Figure \ref{fig:merchant-x1-target} shows the target time-series for feature x1.

%It became clear to us why this feature and many others were being alerted from the start of the dataset until the end. The domain of values is significantly different. The reference period was cleaned of missing values, usually represented by -1. Hence, our reference was built expecting values from 0 to 1 with a certain distribution and in the stream analysis, we found a large volume of missing values or -1. This explains the high number of alerts.

%A follow-up question that is raised is if we should include some mechanism of dealing with missing values in our method. We discard this idea and defend that we should not: missing values are indeed something we want to detect. For instance, if a certain entity responsible for reporting one feature stops working, we want to know that. This scenario is very likely to happen in a distributed Internet-of-Things (IoT) system with hundreds of devices transmitting data.

%missing values streaming

\subsection{Same reference and target periods}
In this next experiment, we set out to test our system in a low drift scenario,...
